---
title: "Midterm_prep"
output: html_document
date: '2023-07-13'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, fig.height = 3)
library(tidyverse)
library(lubridate)
library(readxl)
```


```{r}
mid <- data.frame(pass = c("BK6666.66666+", "O73H89ajhflkah8+", "Jjjjjj56jjjJjk_", "B", "89"))
mid %>% filter(str_length(pass) >= 10, str_detect(pass, "^[:upper:]{1}[^A-Z]"), str_detect(pass, "[:digit:]{2,}"), str_detect(pass, "[\\.,+=-_]$"))

x <- data.frame(X=c(1, 2, 3, 4), Y=c(45, 46, 47, 48))
y <- data.frame(X=c(1, 2, 3, 4, 5, 6), Y=c(45, 46, 47, 48, 50, 60))
#inner_join(x, y)
```


## ggplot

```{r}
ggplot(data = df, mapping = aes(x, y, size, color, group)) +
  geom_histogram(aes(x=duration), 
                 fill="hotpink",  # bar interior
                 color="black",   # bar outline
                 binwidth=7,      # width of the bars
                 boundary=0) +    # boundaries of the bars 
  geom_density(aes(x=duration), # density plot is a smooted histogram
               fill="hotpink",
               color="black") + 
  geom_boxplot(aes(x=as.character(decade),y=duration), # as.character() means treat the column as a string
               fill="red",
               color="blue") +
  geom_violin(aes(x=as.character(decade),y=duration), # density plot for boxplots
              fill="red",
              color="black") +
  geom_point() + # scatterplot
  geom_line() + # lineplot, connect all dots
  geom_smooth(method = "lm", se=FALSE) + # trend line, "lm" means a linear line, se is standard error areas
  geom_hline(yintercept=0) + # horizontal line
  geom_abline(intercept=0, slope=1) +
  geom_bar(aes(fill=method), position=position_dodge2(preserve = "single")) +
  geom_col(aes(fill=method)) +
  
  scale_y_log10() + # log10 scale
  scale_x_log10() +
  scale_y_continuous(trans='log10', breaks = seq(-21,42,by=7),
                     labels = c("Nov 29","Dec 6","Dec 13","Dec 20",
                                "Dec 27","Jan 3","Jan 10","Jan 17",
                                "Jan 24","Jan 31")) +
  facet_wrap(~column) + # create facets based on the column specified
  
  xlab() +
  ylab() +
  labs(color="") +
  ggtitle()
  
```


## dplyr

```{r}
mutate()    # adds new variables that are functions of existing variables
select()    # picks variables based on their names
filter()    # picks cases based on their values
summarise() # reduces multiple values down to a single summary
arrange()   # changes the ordering of the rows

group_by(r) # the variable(s) we are interested in summarizing over
slice_max(data, column, n=5) 
slice_min(data, column, n=5)

planets = read_csv("../../data/exoplanets_default_2021-06-15.csv", skip = 100) %>%
  select(pl_name, discoverymethod, pl_rade, pl_bmasse, disc_year) %>%
  rename(planet = pl_name,
         method = discoverymethod,
         radius = pl_rade,
         mass = pl_bmasse,
         year = disc_year) %>% 
  drop_na()
```

## Manipylating cases

```{r}
filter()  # select rows by a condition
slice()   # select rows by row number
arrange() # order rows
desc()    # used inside of arrange() to specify the order in descending order of a variable
select()  # select variables
          # positive to include; negative to exclude
          # helper functions: starts_with() ends_with() contains() everything() 
mutate()  # add or change one or more variables
transmute() # add or change one or more variable, remove the rest()
rename()  # change the name of a variable
```

## Grouping and Summarizing

```{r}
summarize() # compute a table of summaries
# See the second page of the Cheatsheet for a large vocabulary of summarizing functions
group_by()  # create groups; often used prior to summarize() with the effect that summaries are calculated by group()
```

## Combining Datasets

```{r}
bind_cols() # combine two data frames that have the same observations by binding together columns
bind_rows() # combine two data frame that have the same variables by binding together rows
left_join()
right_join()
inner_join()
full_join() # functions that mutate one or both data frames to join them together

# missing dates
temp = tibble(
  date = seq(ymd("2017-01-01"), ymd("2021-12-31"), 1)
)

temp %>% anti_join(data)
```

```{r}
across()
# select a number of columns with the first argument
# provide a function (or more than one) as the second argument
# apply the function(s) to each selected column
# In the simplest case using across() within mutate(), the columns are replaced by their new values and retain the same names
# We can also use across() within summarize() when we want to create a summary by applying the same function to a selection of columns without needing to list them each individually

where() 
# used inside of across() when we want to use base R functions that return TRUE or FALSE where we would otherwise use tidyverse functions.

pmap_dbl() # will apply some function row-wise and return a numeric valued column of data
# The first argument to pmap_dbl() is a data frame
# As we are not piping into pmap_dbl() but instead call it within mutate(), we use the symbol . to refer to the data frame that was piped into mutate().
```

```{r}
pivot_longer(cols=!zip, names_to = "sex", values_to = "bach_p") 
            # list of columns to pivot, new col name, col name of values
pivot_wider(id_cols=block, names_from=position, values_from=n, values_fill=0)
            # columns to keep, new columns, column taking values, fill empty values with
```


# stringr

```{r}
str_detect(string, pattern)  # detect a pattern in a string
str_count(string, pattern)   # count the number of times a string contains a pattern

str_sub(string, start, end)  # extract a subset by position
str_extract(string, pattern) # extract first match of pattern, NA if no match

str_length(string)           # length of the string
str_pad(string, width, side, pad = " ") # pad strings to a constant width

str_sub(string, start, end) = replacement_string # replace a subset by position
str_replace(string, pattern, replacement) # replace first match with replacement
str_to_lower(string) # change all letters to lower case
str_to_upper(string) # change all letters to upper case

str_c(..., sep = "") # combine multiple strings into one
str_split(string, pattern, n) # splits a string by the pattern
```

Regex:

| — or, match exactly one of the possibilities: can separate more than two chocies
[] — one of, match one of the characters between the brackets
[^]— anything but, match anything not between [^ and ]
ranges — examples: [a-d], [0-5], [a-zA-Z]
() - keep as a group

^ — the start of the string
$ — the end of the string

How often does the previous character/group need to appear?
? — 0 or 1
* — 0 or more
+ — 1 or more
{n} — exactly n times
{n,} — n or more
{m,n} — form m to n times

[:space:] — space, tab, newline
[:blank:] — space and tab
[:punct:] — a collection of punctuation characters
[:symbol:] — a collection of symbols
[:digit:] — a digit
[:lower:] — same as [a-z], a lower-case letter
[:upper:] — same as [A-Z], an upper-case letter
[:alpha:] — same as [a-zA-Z], [:lower:]|[:upper]
[:alnum:] — letter or number, [:alpha:]|[:digit:]
[:graph:] — punctuation, symbols, digits, letters
same as [:punct:]|[:symbol:]|[:alnum:]
. — matches any single character except for a newline




